### 2012-10-17 proposal.tex
Added basic document outline and bibtex placeholder

\documentclass[11pt]{article}
\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array}
\usepackage{bm}
\usepackage{commath}
\usepackage{enumitem}
\usepackage{fancyhdr,fancyvrb}
\usepackage[letterpaper,text={7in,9in},centering]{geometry}
\usepackage[pdftex]{graphicx}
\usepackage{ifthen}
\usepackage{mathrsfs}
\usepackage{multicol}
\usepackage{pifont}
\usepackage{sectsty}
\usepackage{setspace}
\usepackage{stackrel}
\usepackage{stmaryrd}
\usepackage{tensor}
\usepackage{xspace}
\usepackage{natbib}

\pagestyle{fancy}
\fancyhead[RO]{Max Wang \& Edward Gan}

\sectionfont{\large}
\subsectionfont{\normalsize}

\begin{document}
\bibliographystyle{plain}

\title{Proposal: Loose Grammar Inference for Lossy Compression}
\author{Edward Gan \& Max Wang}
\date{10-17-12}
\maketitle

\begin{abstract}
Test (\cite{sequitur})
\end{abstract}

\section{Introduction and Goals}

\section{Background}

\subsection{Grammar Compression}

\subsection{Lossy Compression}

\section{Research Plan}

\subsection{Algorithm Design}

\subsection{Evaluation Methodology}

\subsection{Implementation and Resources}

\subsection{Possible Issues}

\bibliography{sources}

\end{document}