### 2012-10-18 proposal.tex
content vomit pt 1

\documentclass[11pt]{article}
\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array}
\usepackage{bm}
\usepackage{commath}
\usepackage{enumitem}
\usepackage{fancyhdr,fancyvrb}
\usepackage[letterpaper,text={7in,9in},centering]{geometry}
\usepackage[pdftex]{graphicx}
\usepackage{ifthen}
\usepackage{mathrsfs}
\usepackage{multicol}
\usepackage{pifont}
\usepackage{sectsty}
\usepackage{setspace}
\usepackage{stackrel}
\usepackage{stmaryrd}
\usepackage{tensor}
\usepackage{xspace}
\usepackage{natbib}

\pagestyle{fancy}
\fancyhead[RO]{Max Wang \& Edward Gan}

\sectionfont{\large}
\subsectionfont{\normalsize}

\begin{document}
\bibliographystyle{plain}

\title{Proposal: Loose Grammar Inference for Lossy Compression}
\author{Edward Gan \& Max Wang}
\date{10-17-12}
\maketitle

\begin{abstract}
Test (\cite{sequitur})
\end{abstract}

\section{Introduction and Goals}

\section{Background}

Though we have not found very much previous work in the area of lossy grammar based compression,
our proposed research will try to build upon powerful results in grammar based compression and
incorporate some of ideas currently used in lossy compression algorithms.

\subsection{Grammar Compression}

As described in the introduction, the proposed research shares many of its goals with the sequitur
algorithm described in \cite{sequitur}. The authors of \cite{sequitur} demonstrate how 
context free grammars offer a means of simultaneously revealing the structure of
data streams and compressing them effectively. They also develop an effective algorithm
based on iteratively rewriting grammars to keep them small and efficient. Sequitur
yields intriguing grammar based analyses of texts and musical scores, but an inspection of its
results yields that it isn't able to detect and compress larger scale structure because
of small fluctuations that occur in natural streams. This was part of our inspiration for
trying to achieve both better compression and better structural analysis by ignoring
small scale fluctuations.

In \cite{sequitur2}, the detailed compression scheme used in the Sequitur algorithm is described.
By sending the grammar over in an implicit, local-pointer based scheme, they are able to
represent the grammar with very little overhead. A similar scheme should be easily adaptable to any
grammar based compression scheme.

Finally, within the realm of Sequitur based compression, \cite{nevillphd} explores a variety
of ad-hoc extensions to the Sequitur algorithm which improve its compression performance
on structured data. These ranged from introducing domain-specific constraints to their
grammar, adding a few steps of backtracking to their normally greedy grammar formation,
to guessing unifications in attempting to infer recursive grammar rules. Though these
may invalidate some of their theoretical results on the asymptotic performance of
Sequitur, in practice they seem to have worked and add on nicely into the grammar inference
framework.

The intution that grammar inference can support many detailed policies is made more formal
in \cite{grammarcodes}. The authors classify the properties a grammar needs to function
as a good compressor, and moreover give a set of reduction rules for putting grammar into
the appropriate form. Within this context, grammar inference algorithms similar to Sequitur
can be described as simple applications of their reduction rules to different ways
of generating base grammars. Even more examples of the variety of grammar inference schemes
possible under this general framework of reducing grammars can be found in 
\cite{efficientgreedy}

There has also been very interesting work analyzing how grammar based algorithms compare with
the theoretical best lossless grammar for a string, in the line of \cite{approximate}. However,
these analytic bounds for lossless grammars are outside the scope of our proposed research.

In summary, established work on grammar inference algorithms provides a solid set of tools
for experimenting with the kinds of local modifications we propose to make to improve
grammar structure by introducing lossiness.

\subsection{Lossy Compression}

\section{Research Plan}

The proposed research project will involve designing new lossy grammar compression
algorithms, implementing them, and comparing their performance with existing methods.
If time permits we would also like to prove lower bounds on the runtime of our
algorithm and upper bounds on the size of the grammar it generates.

\subsection{Algorithm Design}

Starting with an existing grammar inference algorithm, or perhaps with a
simple variation based
on the reduction rules in \cite{grammarcodes}, there are 3 places we might introduce
lossiness into the system. We could
\begin{itemize}
\item perform text-transforms on the input before applying a 
      grammar inference algorithms
\item modify the grammar inference algorithm so that it generates a lossy
      grammar on the fly
\item take a lossless grammar code and apply lossy transforms to it
\end{itemize}

Of these, the first could provide the biggest potential for taking advantage of
source-specific knowledge of what kinds of data we can ignore. The second
allows the most control over precisely how a grammar can be formed, and would not
be too difficult to build into reduction-rule based frameworks. The third would
be the most general, since it could be applied to any CFG, but may be difficult
to get consistent performance from.

After a lossy CFG is obtained, we plan on using standard methods for encoding the CFG
such as those described in \cite{sequitur2}

\subsection{Evaluation Methodology}

\subsection{Implementation and Resources}

\subsection{Possible Issues}

\nocite{*}

\bibliography{sources}

\end{document}